### 2023-07-21 content/submission/plos_revision1/answer_to_reviewers.md

> Reviewer #2: I have carefully evaluated the manuscript titled "Microtomographic investigation of a large corpus of cichlids". The authors present a substantial collection of tomographic images from cichlids collected in Lake Victoria, Africa. They showcase significant cranial, maxillary, and otolith images from 362 specimens and provide open-source Python notebooks to facilitate the manipulation and extraction of specific parts. I believe that this work represents a significant contribution to the understanding of cichlids and the application of Micro-CT imaging in fish studies. However, I have some suggestions to improve the manuscript:

Thank you very much for the kind words.
As we (now) state in the [imaging section](https://habi.github.io/EAWAG-manuscript/#micro-computed-tomographic-imaging) we acquired 372 single tomographic scans of 133 different specimen (updated since submission of the manuscript, as a rest of fish relevant for the study was added).
This means that each specimen was scanned several times, usually one lower-resolution scan of the full head and two single higher-resolution scans for the oral and pharyngeal jaws.

> The introduction section needs improvement.
> It is too simplistic and lacks informative details about the state of the art and the justification for the work.

We've deliberately kept the introduction rather terse.
The justification of the work lies in the non-destructive imaging of unstained samples.
We have expanded the introduction a bit and added references to manuscripts detailing microtomographic imaging in the life sciences.
As this manuscript is not a review, we kindly refuse to add more information on the background of microtomographic imaging.

@@

> The results of the subsection "Principal components analysis of skull landmarks" should be presented in a more accessible manner.
> Consider providing a tutorial or guide for readers to better understand and apply the analysis.
> As it currently stands, this subsection may be considered irrelevant for the paper.

This subsection is currently being 'expanded' into a self-contained publication.
We deliberately only very briefly mentioned the method here, as the upcoming publication is to be submitted soon.

> The procedure for detecting and extracting otoliths in the subsection "Automatic extraction of otoliths" is not clear.
> Please provide clearer instructions or guidelines for readers using the notebook.

As mentioned above, the Jupyter notebook used to detect the otolith position and to extract the otoliths from the scans covering the whole skull are extensively commented.
We have added more explanatory comments and text to the notebook on GitHub, increasing their 'didactiveness'.

@@

> It would be beneficial to include references in the discussion section to support the arguments and provide a broader context for the findings.

The main scientific 'use case' of the method focuses on the automated, non-destructive extraction of the otolith volume from the tomographic data.
To the best of our knowledge, our manuscript is the first to describe such a method.
We have found two references describing otoliths or otolith structure, both on explanted otoliths.
We have added those to the discussion section.

> Overall, the study seems to have spanned a considerable time frame.
> However, the manuscript's content and structure need improvement to reach a wider audience and effectively communicate the significance of the work.
> The authors have provided a valuable set of open-source Python notebooks, which is commendable and will be highly useful for the scientific community.
> However, the workflow needs to be better explained to enable readers to follow the procedures accurately.

As mentioned above, we have improved both the text and the Jupyter notebooks (linked in the text) with explanatory sentences and comments.

### 2023-07-21 manuscript.tex
Merge pull request #15 from habi/rootstock-2023-07-21 [ci skip]
This build is based on
https://github.com/habi/EAWAG-manuscript/commit/3abb043a6e2aed52fd6af13c2ff321e9a640c139.

This commit was created by the following CI build and job:
https://github.com/habi/EAWAG-manuscript/commit/3abb043a6e2aed52fd6af13c2ff321e9a640c139/checks
https://github.com/habi/EAWAG-manuscript/actions/runs/5623370536

\emph{A preprint of this manuscript is available at bioRχiv with the \href{https://doi.org/10.1101/2023.03.30.534917}{doi:10.1101/2023.03.30.534917}.}
\emph{We submitted \href{https://github.com/habi/EAWAG-manuscript/releases/tag/v1.0}{\texttt{v1.0}} (\href{https://habi.github.io/EAWAG-manuscript/v/e1e2ef76a476174a4115937d77457037ddec95df/}{online version}) of the manuscript to \href{https://journals.plos.org/plosone/}{PLOS ONE}.}
\emph{The version you see here \href{https://github.com/habi/EAWAG-manuscript/compare/v1.0...a690fe23582f53c283f038e12d0f163b7ffa128a}{has been updated} since submission.}

This manuscript
(\href{https://habi.github.io/EAWAG-manuscript/v/a690fe23582f53c283f038e12d0f163b7ffa128a/}{permalink})
was automatically generated
from \href{https://github.com/habi/EAWAG-manuscript/tree/a690fe23582f53c283f038e12d0f163b7ffa128a}{habi/EAWAG-manuscript@a690fe2}
on July 14, 2023.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187} ·
  \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/mastodon.svg}
  \href{https://mastodon.social/@habi}{habi}
  Institute of Anatomy, University of Bern, Baltzerstrasse 2, CH-3012 Bern, Switzerland
\item
  \textbf{Mikki Law}
  \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/mlaw-migalig}{mlaw-migalig}
  Aquatic Ecology and Evolution, Institute of Ecology and Evolution, University of Bern, CH-3012 Bern, Switzerland; Department of Fish Ecology and Evolution, Eawag, Swiss Federal Institute for Aquatic Science and Technology, CH-6047 Kastanienbaum, Switzerland
\item
  \textbf{Kassandra Ford}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-4413-2633}{0000-0002-4413-2633} ·
  \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/kassthefish}{kassthefish}
  Department of Biological Sciences, George Washington University, USA; Department of Fish Ecology and Evolution, Eawag, Swiss Federal Institute for Aquatic Science and Technology, CH-6047 Kastanienbaum, Switzerland; Aquatic Ecology and Evolution, Institute of Ecology and Evolution, University of Bern, CH-3012 Bern, Switzerland
\item
  \textbf{Marcel Häsler}
  \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/mphaesler}{mphaesler}
  Aquatic Ecology and Evolution, Institute of Ecology and Evolution, University of Bern, CH-3012 Bern, Switzerland; Department of Fish Ecology and Evolution, Eawag, Swiss Federal Institute for Aquatic Science and Technology, CH-6047 Kastanienbaum, Switzerland
\item
  \textbf{Ole Seehausen}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0001-6598-1434}{0000-0001-6598-1434} ·
  \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/OleSeehausen}{OleSeehausen}
  Aquatic Ecology and Evolution, Institute of Ecology and Evolution, University of Bern, CH-3012 Bern, Switzerland; Department of Fish Ecology and Evolution, Eawag, Swiss Federal Institute for Aquatic Science and Technology, CH-6047 Kastanienbaum, Switzerland
\item
  \textbf{Ruslan Hlushchuk}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-6722-8996}{0000-0002-6722-8996} ·
  \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/RuslanHlushchuk}{RuslanHlushchuk}
  Institute of Anatomy, University of Bern, Baltzerstrasse 2, CH-3012 Bern, Switzerland
\end{itemize}

@@

Depending on the structures of interest, biomedical samples are often only tomographically scanned after being stained with a contrast agent {[}\protect\hyperlink{ref-pN1S9E5p}{11}{]}, most often employing contrast agents containing heavy metals {[}\protect\hyperlink{ref-kkBa4qNV}{12}{]}.
Due to their inherent contrast difference to the surrounding tissue, the structures of interest we touch upon in this manuscript (teeth and bones, i.e.~jaws and skull) are well visualized in native samples.
Enhancing contrast differences by staining prior to tomographic imaging was thus not necessary to for the samples presented herein.

@@

Since the fish greatly varied in their length (specimen length varied between 6 cm and 18 cm), the voxel sizes of each of the acquired datasets also varies greatly.
We acquired datasets with (isometric) voxel sizes ranging from 3.5 to 50 μm.
Due to the geometric magnification employed by the Bruker SkyScan microCT machines, the voxel size of the resulting dataset is depending on the position of the specimen in the machine.
Natually, the field of view of the resulting dataset also depends on the chosen voxel size.
The chosen voxel size is thus scan-specific.
It was always chosen in a way that the region of interest to be images fit into the available lateral field of view of the microCT machine at the smallest necessary voxel size to resolve the structures of interest.
Since the region of interest (either oral or pharyngeal jaws or the complete skulls of the fish) often did not fit into a single field of view \emph{along} the anteroposterior axis of the fish, we often performed several tomographic scans along the rotation axis of the specimen in the machine (since the specimen were all scanned rotating around the anteroposterior axis).
The projection images from those so-called stacked scans are subsequently automatically merged to one stack of reconstructions images by the reconstruction software.

@@

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
Author
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Contributions
\end{minipage} \\
\midrule
\endhead
David Haberthür & Conceptualization, Data curation, Formal Analysis, Investigation, Methodology, Project administration, Software, Visualization, Writing -- original draft, Writing -- review \& editing \\
Mikki Law & Data curation, Investigation, Methodology, Project administration, Visualization, Writing -- review \& editing \\
Kassandra Ford & Investigation, Methodology, Visualization, Writing -- review \& editing \\
Marcel Häsler & Investigation, Project administration, Resources, Writing -- review \& editing \\
Ole Seehausen & Conceptualization, Investigation, Resources, Supervision, Writing -- review \& editing \\
Ruslan Hlushchuk & Conceptualization, Resources, Supervision, Writing -- review \& editing \\
\bottomrule
\end{longtable}

@@

We thank Salome Mwaiko for taking care of the fish collection at Eawag and Mark Charran for helping to find suitable specimens to represent each species.
We are grateful to the \href{https//www.mic.unibe.ch/}{Microscopy Imaging Center} of the University of Bern for the infrastructural support.
We thank the \texttt{manubot} project {[}\protect\hyperlink{ref-YuJbg3zO}{36}{]} for helping us write this manuscript collaboratively.

@@

The three-dimensional view of sample 104016 was generated in \href{https://github.com/habi/EAWAG/blob/master/ExtractOtoliths.ipynb}{the otolith extraction notebook} and saved as a \href{https://github.com/habi/EAWAG-manuscript/blob/main/content/data/104016.head.rec.otolith.region.3D.html}{self-contained \texttt{HTML} file} with \href{https://github.com/K3D-tools/K3D-jupyter}{\emph{K3D-jupyter}}.
A copy of this \texttt{HTML} file can be viewed and interacted with through the \href{https://htmlpreview.github.io/?https://github.com/habi/EAWAG-manuscript/blob/main/content/data/104016.head.rec.otolith.region.3D.html}{GitHub HTML preview}.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-TUQFenhc}{}}%
\CSLLeftMargin{1. }
\CSLRightInline{Kocher TD. Adaptive evolution and explosive speciation: the cichlid fish model. Nat Rev Genet. 2004;5: 288--298. doi:\href{https://doi.org/10.1038/nrg1316}{10.1038/nrg1316}}

\leavevmode\vadjust pre{\hypertarget{ref-W2f2SDjn}{}}%
\CSLLeftMargin{2. }
\CSLRightInline{Seehausen O. African cichlid fish: a model system in adaptive radiation research. Proc R Soc B. 2006;273: 1987--1998. doi:\href{https://doi.org/10.1098/rspb.2006.3539}{10.1098/rspb.2006.3539}}

\leavevmode\vadjust pre{\hypertarget{ref-WVIa4CH0}{}}%
\CSLLeftMargin{3. }
\CSLRightInline{Konings AF, Wisor JM, Stauffer JR Jr. Microcomputed tomography used to link head morphology and observed feeding behavior in cichlids of Lake Malaŵi. Ecol Evol. 2021;11: 4605--4615. doi:\href{https://doi.org/10.1002/ece3.7359}{10.1002/ece3.7359}}

\leavevmode\vadjust pre{\hypertarget{ref-1GEljkDsf}{}}%
\CSLLeftMargin{4. }
\CSLRightInline{Liem KF. Evolutionary Strategies and Morphological Innovations: Cichlid Pharyngeal Jaws. Systematic Zoology. 1973;22: 425. doi:\href{https://doi.org/10.2307/2412950}{10.2307/2412950}}

\leavevmode\vadjust pre{\hypertarget{ref-c3poTogo}{}}%
\CSLLeftMargin{5. }
\CSLRightInline{Seehausen O. Process and pattern in cichlid radiations -- inferences for understanding unusually high rates of evolutionary diversification. New Phytologist. 2015;207: 304--312. doi:\href{https://doi.org/10.1111/nph.13450}{10.1111/nph.13450}}

\leavevmode\vadjust pre{\hypertarget{ref-19h9vxsYG}{}}%
\CSLLeftMargin{6. }
\CSLRightInline{Ford KL, Albert JS, Summers AP, Hedrick BP, Schachner ER, Jones AS, et al. A New Era of Morphological Investigations: Reviewing Methods for Comparative Anatomical Studies. Integrative Organismal Biology. 2023;5. doi:\href{https://doi.org/10.1093/iob/obad008}{10.1093/iob/obad008}}

\leavevmode\vadjust pre{\hypertarget{ref-IKVBgaCE}{}}%
\CSLLeftMargin{7. }
\CSLRightInline{Rawson SD, Maksimcuka J, Withers PJ, Cartmell SH. X-ray computed tomography in life sciences. BMC Biol. 2020;18. doi:\href{https://doi.org/10.1186/s12915-020-0753-2}{10.1186/s12915-020-0753-2}}

\leavevmode\vadjust pre{\hypertarget{ref-8Pmc2mI8}{}}%
\CSLLeftMargin{8. }
\CSLRightInline{Messerli M, Aaldijk D, Haberthür D, Röss H, García-Poyatos C, Sande-Melón M, et al. Adaptation mechanism of the adult zebrafish respiratory organ to endurance training. Shelden EA, editor. PLoS ONE. 2020;15: e0228333. doi:\href{https://doi.org/10.1371/journal.pone.0228333}{10.1371/journal.pone.0228333}}

\leavevmode\vadjust pre{\hypertarget{ref-BDrCSu8p}{}}%
\CSLLeftMargin{9. }
\CSLRightInline{Ford KL, Peterson R, Bernt M, Albert JS. Convergence is Only Skin Deep: Craniofacial Evolution in Electric Fishes from South America and Africa (Apteronotidae and Mormyridae). Integrative Organismal Biology. 2022;4. doi:\href{https://doi.org/10.1093/iob/obac022}{10.1093/iob/obac022}}

\leavevmode\vadjust pre{\hypertarget{ref-3rxGCEvJ}{}}%
\CSLLeftMargin{10. }
\CSLRightInline{Available: \url{https://osf.io/ecmz4}}

\leavevmode\vadjust pre{\hypertarget{ref-pN1S9E5p}{}}%
\CSLLeftMargin{11. }
\CSLRightInline{Descamps E, Sochacka A, De Kegel B, Van Loo D, Van Hoorebeke L, Adriaens D. Soft tissue discrimination with contrast agents using micro-CT scanning. Belg J Zool. 2020;144. doi:\href{https://doi.org/10.26496/bjz.2014.63}{10.26496/bjz.2014.63}}

\leavevmode\vadjust pre{\hypertarget{ref-kkBa4qNV}{}}%
\CSLLeftMargin{12. }
\CSLRightInline{Babaei F, Hong TLC, Yeung K, Cheng SH, Lam YW. Contrast-Enhanced X-Ray Micro-Computed Tomography as a Versatile Method for Anatomical Studies of Adult Zebrafish. Zebrafish. 2016;13: 310--316. doi:\href{https://doi.org/10.1089/zeb.2016.1245}{10.1089/zeb.2016.1245}}

\leavevmode\vadjust pre{\hypertarget{ref-VCZPOv2f}{}}%
\CSLLeftMargin{13. }
\CSLRightInline{Available: \url{https://github.com/TomoGraphics/Hol3Drs/blob/master/STL/EAWAG.Fish.stl}}

\leavevmode\vadjust pre{\hypertarget{ref-115PPSuQp}{}}%
\CSLLeftMargin{14. }
\CSLRightInline{Haberthür D. TomoGraphics/Hol3Drs: A release. Zenodo; 2019. doi:\href{https://doi.org/10.5281/zenodo.2587555}{10.5281/zenodo.2587555}}

\leavevmode\vadjust pre{\hypertarget{ref-wsjdcTeC}{}}%
\CSLLeftMargin{15. }
\CSLRightInline{OpenSCAD. {[}cited 14 Jul 2023{]}. Available: \url{https://openscad.org}}

\leavevmode\vadjust pre{\hypertarget{ref-f7OzrzUv}{}}%
\CSLLeftMargin{16. }
\CSLRightInline{Available: \url{https://github.com/TomoGraphics/Hol3Drs/blob/master/EAWAG.Fish.scad}}

\leavevmode\vadjust pre{\hypertarget{ref-ZjgrPHTa}{}}%
\CSLLeftMargin{17. }
\CSLRightInline{Available: \url{https://github.com/habi/EAWAG/blob/master/rsync-fishes.sh}}

\leavevmode\vadjust pre{\hypertarget{ref-pQ6Wbz73}{}}%
\CSLLeftMargin{18. }
\CSLRightInline{Kluyver T, Ragan-Kelley B, Pérez F, Granger B, Bussonnier M, Frederic J, et al. Jupyter Notebooks -- a publishing format for reproducible computational workflows. In: Loizides F, Scmidt B, editors. IOS Press; 2016. pp. 87--90. doi:\href{https://doi.org/10.3233/978-1-61499-649-1-87}{10.3233/978-1-61499-649-1-87}}

>>>>

\emph{A preprint of this manuscript is available at bioRχiv with the \href{https://doi.org/10.1101/2023.03.30.534917}{doi:10.1101/2023.03.30.534917}.}
\emph{We submitted \href{https://github.com/habi/EAWAG-manuscript/releases/tag/v1.0}{\texttt{v1.0}} (\href{https://habi.github.io/EAWAG-manuscript/v/e1e2ef76a476174a4115937d77457037ddec95df/}{archived here}) of the manuscript to \href{https://journals.plos.org/plosone/}{PLOS ONE}.}
\emph{A revised version \href{https://github.com/habi/EAWAG-manuscript/releases/tag/v1.1}{\texttt{v1.1}} (\href{https://habi.github.io/EAWAG-manuscript/v/c49d0b8ac8f078c226318972b06e8c4a39100bb2/}{archived here}) has been submitted.}
\emph{The version you see here \href{https://github.com/habi/EAWAG-manuscript/compare/v1.1...3abb043a6e2aed52fd6af13c2ff321e9a640c139\#files_bucket}{is further updated} since re-submission.}

This manuscript
(\href{https://habi.github.io/EAWAG-manuscript/v/3abb043a6e2aed52fd6af13c2ff321e9a640c139/}{permalink})
was automatically generated
from \href{https://github.com/habi/EAWAG-manuscript/tree/3abb043a6e2aed52fd6af13c2ff321e9a640c139}{habi/EAWAG-manuscript@3abb043}
on July 21, 2023.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187} ·
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/mastodon.svg}
  \href{https://mastodon.social/@habi}{@habi@mastodon.social}
  Institute of Anatomy, University of Bern, Baltzerstrasse 2, CH-3012 Bern, Switzerland
\item
  \textbf{Mikki Law}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/mlaw-migalig}{mlaw-migalig}
  Aquatic Ecology and Evolution, Institute of Ecology and Evolution, University of Bern, CH-3012 Bern, Switzerland; Department of Fish Ecology and Evolution, Eawag, Swiss Federal Institute for Aquatic Science and Technology, CH-6047 Kastanienbaum, Switzerland
\item
  \textbf{Kassandra Ford}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-4413-2633}{0000-0002-4413-2633} ·
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/kassthefish}{kassthefish}
  Department of Biological Sciences, George Washington University, USA; Department of Fish Ecology and Evolution, Eawag, Swiss Federal Institute for Aquatic Science and Technology, CH-6047 Kastanienbaum, Switzerland; Aquatic Ecology and Evolution, Institute of Ecology and Evolution, University of Bern, CH-3012 Bern, Switzerland
\item
  \textbf{Marcel Häsler}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/mphaesler}{mphaesler}
  Aquatic Ecology and Evolution, Institute of Ecology and Evolution, University of Bern, CH-3012 Bern, Switzerland; Department of Fish Ecology and Evolution, Eawag, Swiss Federal Institute for Aquatic Science and Technology, CH-6047 Kastanienbaum, Switzerland
\item
  \textbf{Ole Seehausen}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0001-6598-1434}{0000-0001-6598-1434} ·
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/OleSeehausen}{OleSeehausen}
  Aquatic Ecology and Evolution, Institute of Ecology and Evolution, University of Bern, CH-3012 Bern, Switzerland; Department of Fish Ecology and Evolution, Eawag, Swiss Federal Institute for Aquatic Science and Technology, CH-6047 Kastanienbaum, Switzerland
\item
  \textbf{Ruslan Hlushchuk}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-6722-8996}{0000-0002-6722-8996} ·
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/RuslanHlushchuk}{RuslanHlushchuk}
  Institute of Anatomy, University of Bern, Baltzerstrasse 2, CH-3012 Bern, Switzerland
\end{itemize}

\leavevmode\vadjust pre{\hypertarget{correspondence}{}}%
✉ --- Correspondence possible via \href{https://github.com/habi/EAWAG-manuscript/issues}{GitHub Issues}

@@

Depending on the structures of interest to be assessed, biomedical samples are often only tomographically scanned after them being stained with a contrast agent {[}\protect\hyperlink{ref-pN1S9E5p}{11}{]}, most often employing contrast agents containing heavy metals {[}\protect\hyperlink{ref-kkBa4qNV}{12}{]}.
This is done to increase the contrast difference between different structures in the sample.
Due to their inherent contrast difference to the surrounding tissue, the structures of interest we touch upon in this manuscript (teeth and bones, i.e.~jaws and skull) are well visualized in native samples.
Enhancing contrast differences by staining prior to tomographic imaging was thus not necessary to for the samples presented herein.

@@

Since the fish greatly varied in their length (specimen length varied between 6 cm and 18 cm), the voxel sizes of each of the acquired datasets also varies greatly.
We acquired datasets with (isometric) voxel sizes ranging from 3.5 to 50 μm.
Due to the geometric magnification employed by the Bruker SkyScan microCT machines, the voxel size of the resulting dataset is depending on the position of the specimen in the machine.
Naturally, the field of view of the resulting dataset also depends on the chosen voxel size.
The chosen voxel size is thus scan-specific.
It was always chosen in a way that the region of interest to be images fit into the available lateral field of view of the microCT machine at the smallest necessary voxel size to resolve the structures of interest.
Since the region of interest (either oral or pharyngeal jaws or the complete skulls of the fish) often did not fit into a single field of view \emph{along} the anteroposterior axis of the fish, we often performed several tomographic scans along the rotation axis of the specimen in the machine (since the specimen were all scanned rotating around the anteroposterior axis).
The projection images from those so-called stacked scans are subsequently automatically merged to one stack of reconstructions images by the reconstruction software.

@@

To our best knowledge, no fully automated method for non-destructive otolith extraction from high-resolution tomographic data has been published up to now.
Vasconcelos-Filho et al. {[}\protect\hyperlink{ref-PgMc4mCy}{36}{]} show a potentially objective method to count growth rings in otoliths of six fish species, but performed their study on extracted otoliths ``embedded in a Styrofoam cube'', while our method non-destructively extracts the otoliths from the tomographic data.
Gu et al. {[}\protect\hyperlink{ref-HYXdLQTV}{37}{]} also provide detailed insight into otolith and combine microtomographic imaging (with unknown voxel size) with scanning electron microscopy, but also mention that the ``otoliths were removed {[}from \emph{Bahaba taipingensis}{]}, cleaned and kept dry for morphologic observation''.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Author
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Contributions
\end{minipage} \\
\midrule()
\endhead
David Haberthür & Conceptualization, Data curation, Formal Analysis, Investigation, Methodology, Project administration, Software, Visualization, Writing -- original draft, Writing -- review \& editing \\
Mikki Law & Data curation, Investigation, Methodology, Project administration, Visualization, Writing -- review \& editing \\
Kassandra Ford & Investigation, Methodology, Visualization, Writing -- review \& editing \\
Marcel Häsler & Investigation, Project administration, Resources, Writing -- review \& editing \\
Ole Seehausen & Conceptualization, Investigation, Resources, Supervision, Writing -- review \& editing \\
Ruslan Hlushchuk & Conceptualization, Resources, Supervision, Writing -- review \& editing \\
\bottomrule()
\end{longtable}

@@

We thank Salome Mwaiko for taking care of the fish collection at Eawag and Mark Charran for helping to find suitable specimens to represent each species.
We are grateful to the \href{https//www.mic.unibe.ch/}{Microscopy Imaging Center} of the University of Bern for the infrastructural support.
We thank the \texttt{manubot} project {[}\protect\hyperlink{ref-YuJbg3zO}{38}{]} for helping us write this manuscript collaboratively.

@@

The three-dimensional view of sample 104016 was generated in \href{https://github.com/habi/EAWAG/blob/master/ExtractOtoliths.ipynb}{the otolith extraction notebook} and saved as a \href{https://github.com/habi/EAWAG-manuscript/blob/main/content/data/104016.head.rec.otolith.region.3D.html}{self-contained \texttt{HTML} file} with \href{https://github.com/K3D-tools/K3D-jupyter}{\emph{K3D-jupyter}}.
A copy of this \texttt{HTML} file can be viewed and interacted with through the \href{https://htmlpreview.github.io/?https://github.com/habi/EAWAG-manuscript/blob/main/content/data/104016_Enterochromis_I_cinctus_St_E.head.rec.Otolith.Region.3D.html}{GitHub HTML preview}.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-TUQFenhc}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{Kocher TD. Adaptive evolution and explosive speciation: the cichlid fish model. Nat Rev Genet. 2004;5: 288--298. doi:\href{https://doi.org/10.1038/nrg1316}{10.1038/nrg1316}}

\leavevmode\vadjust pre{\hypertarget{ref-W2f2SDjn}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{Seehausen O. African cichlid fish: a model system in adaptive radiation research. Proc R Soc B. 2006;273: 1987--1998. doi:\href{https://doi.org/10.1098/rspb.2006.3539}{10.1098/rspb.2006.3539}}

\leavevmode\vadjust pre{\hypertarget{ref-WVIa4CH0}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{Konings AF, Wisor JM, Stauffer JR Jr. Microcomputed tomography used to link head morphology and observed feeding behavior in cichlids of Lake Malaŵi. Ecol Evol. 2021;11: 4605--4615. doi:\href{https://doi.org/10.1002/ece3.7359}{10.1002/ece3.7359}}

\leavevmode\vadjust pre{\hypertarget{ref-1GEljkDsf}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{Liem KF. Evolutionary Strategies and Morphological Innovations: Cichlid Pharyngeal Jaws. Systematic Zoology. 1973;22: 425. doi:\href{https://doi.org/10.2307/2412950}{10.2307/2412950}}

\leavevmode\vadjust pre{\hypertarget{ref-c3poTogo}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{Seehausen O. Process and pattern in cichlid radiations -- inferences for understanding unusually high rates of evolutionary diversification. New Phytologist. 2015;207: 304--312. doi:\href{https://doi.org/10.1111/nph.13450}{10.1111/nph.13450}}

\leavevmode\vadjust pre{\hypertarget{ref-19h9vxsYG}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{Ford KL, Albert JS, Summers AP, Hedrick BP, Schachner ER, Jones AS, et al. A New Era of Morphological Investigations: Reviewing Methods for Comparative Anatomical Studies. Integrative Organismal Biology. 2023;5. doi:\href{https://doi.org/10.1093/iob/obad008}{10.1093/iob/obad008}}

\leavevmode\vadjust pre{\hypertarget{ref-IKVBgaCE}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{Rawson SD, Maksimcuka J, Withers PJ, Cartmell SH. X-ray computed tomography in life sciences. BMC Biol. 2020;18. doi:\href{https://doi.org/10.1186/s12915-020-0753-2}{10.1186/s12915-020-0753-2}}

\leavevmode\vadjust pre{\hypertarget{ref-8Pmc2mI8}{}}%
\CSLLeftMargin{8. }%
\CSLRightInline{Messerli M, Aaldijk D, Haberthür D, Röss H, García-Poyatos C, Sande-Melón M, et al. Adaptation mechanism of the adult zebrafish respiratory organ to endurance training. Shelden EA, editor. PLoS ONE. 2020;15: e0228333. doi:\href{https://doi.org/10.1371/journal.pone.0228333}{10.1371/journal.pone.0228333}}

\leavevmode\vadjust pre{\hypertarget{ref-BDrCSu8p}{}}%
\CSLLeftMargin{9. }%
\CSLRightInline{Ford KL, Peterson R, Bernt M, Albert JS. Convergence is Only Skin Deep: Craniofacial Evolution in Electric Fishes from South America and Africa (Apteronotidae and Mormyridae). Integrative Organismal Biology. 2022;4. doi:\href{https://doi.org/10.1093/iob/obac022}{10.1093/iob/obac022}}

\leavevmode\vadjust pre{\hypertarget{ref-3rxGCEvJ}{}}%
\CSLLeftMargin{10. }%
\CSLRightInline{Available: \url{https://osf.io/ecmz4}}

\leavevmode\vadjust pre{\hypertarget{ref-pN1S9E5p}{}}%
\CSLLeftMargin{11. }%
\CSLRightInline{Descamps E, Sochacka A, De Kegel B, Van Loo D, Van Hoorebeke L, Adriaens D. Soft tissue discrimination with contrast agents using micro-CT scanning. Belg J Zool. 2020;144. doi:\href{https://doi.org/10.26496/bjz.2014.63}{10.26496/bjz.2014.63}}

\leavevmode\vadjust pre{\hypertarget{ref-kkBa4qNV}{}}%
\CSLLeftMargin{12. }%
\CSLRightInline{Babaei F, Hong TLC, Yeung K, Cheng SH, Lam YW. Contrast-Enhanced X-Ray Micro-Computed Tomography as a Versatile Method for Anatomical Studies of Adult Zebrafish. Zebrafish. 2016;13: 310--316. doi:\href{https://doi.org/10.1089/zeb.2016.1245}{10.1089/zeb.2016.1245}}

\leavevmode\vadjust pre{\hypertarget{ref-VCZPOv2f}{}}%
\CSLLeftMargin{13. }%
\CSLRightInline{Available: \url{https://github.com/TomoGraphics/Hol3Drs/blob/master/STL/EAWAG.Fish.stl}}

\leavevmode\vadjust pre{\hypertarget{ref-115PPSuQp}{}}%
\CSLLeftMargin{14. }%
\CSLRightInline{Haberthür D. TomoGraphics/Hol3Drs: A release. Zenodo; 2019. doi:\href{https://doi.org/10.5281/zenodo.2587555}{10.5281/zenodo.2587555}}

\leavevmode\vadjust pre{\hypertarget{ref-wsjdcTeC}{}}%
\CSLLeftMargin{15. }%
\CSLRightInline{OpenSCAD. {[}cited 21 Jul 2023{]}. Available: \url{https://openscad.org}}

\leavevmode\vadjust pre{\hypertarget{ref-f7OzrzUv}{}}%
\CSLLeftMargin{16. }%
\CSLRightInline{Available: \url{https://github.com/TomoGraphics/Hol3Drs/blob/master/EAWAG.Fish.scad}}

\leavevmode\vadjust pre{\hypertarget{ref-ZjgrPHTa}{}}%
\CSLLeftMargin{17. }%
\CSLRightInline{Available: \url{https://github.com/habi/EAWAG/blob/master/rsync-fishes.sh}}

\leavevmode\vadjust pre{\hypertarget{ref-pQ6Wbz73}{}}%
\CSLLeftMargin{18. }%
\CSLRightInline{Kluyver T, Ragan-Kelley B, Pérez F, Granger B, Bussonnier M, Frederic J, et al. Jupyter Notebooks -- a publishing format for reproducible computational workflows. In: Loizides F, Scmidt B, editors. IOS Press; 2016. pp. 87--90. doi:\href{https://doi.org/10.3233/978-1-61499-649-1-87}{10.3233/978-1-61499-649-1-87}}

\leavevmode\vadjust pre{\hypertarget{ref-1HteOscVd}{}}%
\CSLLeftMargin{19. }%
\CSLRightInline{Haberthür D. habi/EAWAG: Jupyter notebooks for preparing tomographic scans of a collection of cichlids for analysis. Zenodo; 2022. doi:\href{https://doi.org/10.5281/zenodo.6798632}{10.5281/zenodo.6798632}}
